---
title: "Tutorial for TRexDAD Package"
author: "Yunyi Cheng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial for TRexDAD Package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
# Introduction
The TRexDAD package offers tools for optimizing gene editing designs. This vignette demonstrates the use of key functions in the package.

## Installation

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

```{r setup}
require("devtools")
# Install from GitHub
devtools::install_github("yunyicheng/TRexDAD", build_vignettes = TRUE)
# Load the package
library(TRexDAD)
```

To list all functions available in the package:
```
```{r}
lsf.str("package:TRexDAD")
```

To list all sample datasets available in the package:
``` r
data(package = "DeregGenes")
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
require("devtools")
devtools::install_github("yunyicheng/TRexDAD")
library(TRexDAD)
```
